---
title: "AD timeline paper"
output:
  html_document:
    df_print: paged
---



```{r header, include=FALSE}
library(here)
source( here::here("scripts", "00-libraries.R") )
source( here::here("scripts", "01-function varfile.R") )
source( here::here("scripts", "01-function plotloop.R") )
source( here::here("scripts", "01-function mortplot.R") )
source( here::here("scripts", "01-function nacctable.R") )
source( here::here("scripts", "01-function nacctable2.R") )
source( here::here("scripts", "02-import data.R") )


```
## Overview

Describe and model timeline for each CDR stage of AD and identify factors that mediate the timeline.


```{r parms, include=FALSE}
col.array<-c("forestgreen","navy","black")
left.table=-1
table.shift=1
table.shift2=.5
Y.line=0
x.cross=0
x.upper=10
x.lower=0
y.lower=0
y.upper=1
tic.int=2
Ylab="Percent"
x.text.pos=(-1)
p.lim=0.25 # width of progression length CI


Ns.pos=-0.2
```


## NACC data

Data construction (from cdr=X):\
1. select all NACC participants with cdr=x-1 at some point\
2. find first cdr=x--define this as first visit\
3. events defined in this order:\
+ event = 1, time to first CDR $\ge$ x+1\
+ event = 2, time to death (in the case that CDR never gets past x), only coded when CDR evaluated within 1.5y of death\
+ event = 0, time to last CDR evaluation


### CDR 0.5 progression (incident CDR$=0.5$ to CDR $\ge 2$)

```{r cdr05, echo=FALSE}
mat1=nacc.table(cdr05.1)
options(knitr.kable.NA = '')
kable(mat1,caption="",align="llcclcclccl")
```

mean years observed (SD)

```{r cdr5.yrs, echo=FALSE}
y=rep(0,4)
dat=cdr05.1
y[1]=round(mean(dat$yrs),1)
y[2]=round(mean(dat$yrs[dat$event==0]),1)
y[3]=round(mean(dat$yrs[dat$event==1]),1)
y[4]=round(mean(dat$yrs[dat$event==2]),1)

y2=rep(0,4)
y2[1]=round(sqrt(var(dat$yrs)),1)
y2[2]=round(sqrt(var(dat$yrs[dat$event==0])),1)
y2[3]=round(sqrt(var(dat$yrs[dat$event==1])),1)
y2[4]=round(sqrt(var(dat$yrs[dat$event==2])),1)

a=c(paste("All: ",y[1]," (",y2[1],")",sep=""),
paste("Censored: ",y[2]," (",y2[2],")",sep=""),
paste("AD progression: ",y[3]," (",y2[3],")",sep=""),
paste("Deceased: ",y[4]," (",y2[4],")",sep=""))

kable(a,caption="")

```

```{r CDR.5, echo=FALSE}
################################################

main.text0="CDR 0.5 AD to >=1"
x.text="years from CDR 0.5 AD"
dat0=cdr05.1
dat0=dat0[dat0$nacc_addx_t1==1,]


age.table=table(dat0$age.group10)
age.groups=names(age.table)
t.max=10
i=1
main.text=paste("All ages: ",main.text0,sep="")
  dat1=dat0
  plot.loop()
for(i in 1:length(age.groups)){
  main.text=paste("Age ",age.groups[i],": ",main.text0,sep="")
  dat1=dat0[dat0$age.group10==age.groups[i],]
  plot.loop()}



```
Look at 2, 5 and 10y outcomes (where death trumps AD)

```{r CDR.5.outcomes, echo=FALSE}
################################################
#CDR 0.5 AD to >=1

dat=cdr05.1
table(dat$age.group10)

t1=2
out.2y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.2y=round(proportions(table(out.2y,dat$age.group10,useNA="ifany"),margin=2),2)
table.2y

t1=5
y1=ifelse(dat$yrs>t1,1,0)
y2=ifelse(dat$yrs_death>t1,1,0)

out.5y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                    ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))

table.5y=round(proportions(table(out.5y,dat$age.group10,useNA="ifany"),margin=2),2)
table.5y

t1=10
out.10y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.10y=round(proportions(table(out.10y,dat$age.group10,useNA="ifany"),margin=2),2)
table.10y
```



### CDR 1 progression (incident CDR$=1$ to CDR $\ge 2$)

```{r cdr1, echo=FALSE}
mat1=nacc.table(cdr1.23)
options(knitr.kable.NA = '')
kable(mat1,caption="",align="llcclcclccl")
```

mean years observed (SD)

```{r cdr1.yrs, echo=FALSE}
y=rep(0,4)
dat=cdr1.23
y[1]=round(mean(dat$yrs),1)
y[2]=round(mean(dat$yrs[dat$event==0]),1)
y[3]=round(mean(dat$yrs[dat$event==1]),1)
y[4]=round(mean(dat$yrs[dat$event==2]),1)

y2=rep(0,4)
y2[1]=round(sqrt(var(dat$yrs)),1)
y2[2]=round(sqrt(var(dat$yrs[dat$event==0])),1)
y2[3]=round(sqrt(var(dat$yrs[dat$event==1])),1)
y2[4]=round(sqrt(var(dat$yrs[dat$event==2])),1)

a=c(paste("All: ",y[1]," (",y2[1],")",sep=""),
paste("Censored: ",y[2]," (",y2[2],")",sep=""),
paste("AD progression: ",y[3]," (",y2[3],")",sep=""),
paste("Deceased: ",y[4]," (",y2[4],")",sep=""))

kable(a,caption="")

```
```{r CDR.1, echo=FALSE}
################################################
main.text0="CDR 1 AD to >=2"
x.text="years from CDR 1 AD"
dat0=cdr1.23

main.text=paste("All ages: ",main.text0,sep="")
dat1=dat0
plot.loop()
```

```{r CDR.1.age, echo=FALSE}
################################################  
age.table=table(dat0$age.group10)
age.groups=names(age.table)
t.max=8

par(mfrow=c(1,1))
for(i in 1:length(age.groups)){
  main.text=paste("Age ",age.groups[i],": ",main.text0,sep="")
  dat1=dat0[dat0$age.group10==age.groups[i],]
  plot.loop()}
```
CDR 1 AD outcomes

```{r CDR.1.outcomes, echo=FALSE}
################################################
dat=cdr1.23
table(dat$age.group10)

t1=2
out.2y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.2y=round(proportions(table(out.2y,dat$age.group10,useNA="ifany"),margin=2),2)
table.2y

t1=5
y1=ifelse(dat$yrs>t1,1,0)
y2=ifelse(dat$yrs_death>t1,1,0)

out.5y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))

table.5y=round(proportions(table(out.5y,dat$age.group10,useNA="ifany"),margin=2),2)
table.5y

t1=10
out.10y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
               ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                      ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.10y=round(proportions(table(out.10y,dat$age.group10,useNA="ifany"),margin=2),2)
table.10y
```

### CDR 2 progression (incident CDR$=2$ to CDR $=3$)

```{r cdr2, echo=FALSE}
mat1=nacc.table(cdr2.3)
options(knitr.kable.NA = '')
kable(mat1,caption="",align="llcclcclccl")
```

mean years observed (SD)

```{r cdr2.yrs, echo=FALSE}
y=rep(0,4)
dat=cdr2.3
y[1]=round(mean(dat$yrs),1)
y[2]=round(mean(dat$yrs[dat$event==0]),1)
y[3]=round(mean(dat$yrs[dat$event==1]),1)
y[4]=round(mean(dat$yrs[dat$event==2]),1)

y2=rep(0,4)
y2[1]=round(sqrt(var(dat$yrs)),1)
y2[2]=round(sqrt(var(dat$yrs[dat$event==0])),1)
y2[3]=round(sqrt(var(dat$yrs[dat$event==1])),1)
y2[4]=round(sqrt(var(dat$yrs[dat$event==2])),1)

a=c(paste("All: ",y[1]," (",y2[1],")",sep=""),
paste("Censored: ",y[2]," (",y2[2],")",sep=""),
paste("AD progression: ",y[3]," (",y2[3],")",sep=""),
paste("Deceased: ",y[4]," (",y2[4],")",sep=""))

kable(a,caption="")

```
```{r CDR.2, echo=FALSE}
################################################
main.text0="CDR 2 AD to 3"
x.text="years from CDR 2 AD"
dat0=cdr2.3

main.text=paste("All ages: ",main.text0,sep="")
dat1=dat0
plot.loop()


```

```{r CDR.2.age, echo=FALSE}
################################################
age.table=table(dat0$age.group10)
age.groups=names(age.table)
t.max=8

par(mfrow=c(1,1))
for(i in 1:length(age.groups)){
  main.text=paste("Age ",age.groups[i],": ",main.text0,sep="")
  dat1=dat0[dat0$age.group10==age.groups[i],]
  plot.loop()}

```
CDR 2 AD outcomes

```{r CDR.2.outcomes, echo=FALSE}
################################################
dat=cdr2.3
table(dat$age.group10)

t1=2
out.2y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.2y=round(proportions(table(out.2y,dat$age.group10,useNA="ifany"),margin=2),2)
table.2y


t1=5
y1=ifelse(dat$yrs>t1,1,0)
y2=ifelse(dat$yrs_death>t1,1,0)

out.5y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
              ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                     ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))

table.5y=round(proportions(table(out.5y,dat$age.group10,useNA="ifany"),margin=2),2)
table.5y

t1=10
out.10y=ifelse((dat$yrs_death<=t1)&(dat$death_event==1),"2-Deceased",
               ifelse((dat$event==1)&(dat$yrs<=t1),"1-AD progression",
                      ifelse(dat$yrs>t1,"0-Alive w/o progression", NA)))
table.10y=round(proportions(table(out.10y,dat$age.group10,useNA="ifany"),margin=2),2)
table.10y
```

### CDR 3 progression (incident CDR$=3$ to death)

```{r cdr3, echo=FALSE}
mat1=nacc.table2(cdr3.death)
options(knitr.kable.NA = '')
kable(mat1,caption="",align="llcclcclccl")
```

mean years observed (SD)

```{r cdr3.yrs, echo=FALSE}
y=rep(0,3)
dat=cdr3.death
y[1]=round(mean(dat$yrs),1)
y[2]=round(mean(dat$yrs[dat$event==0]),1)
y[3]=round(mean(dat$yrs[dat$event==1]),1)

y2=rep(0,3)
y2[1]=round(sqrt(var(dat$yrs)),1)
y2[2]=round(sqrt(var(dat$yrs[dat$event==0])),1)
y2[3]=round(sqrt(var(dat$yrs[dat$event==1])),1)

a=c(paste("All: ",y[1]," (",y2[1],")",sep=""),
paste("Censored: ",y[2]," (",y2[2],")",sep=""),
paste("Deceased: ",y[3]," (",y2[3],")",sep=""))

kable(a,caption="")

```



```{r CDR.3, echo=FALSE}
################################################
main.text0="CDR 3 AD mortality"
x.text="years from CDR 3 AD"
dat0=cdr3.death

main.text=paste("All ages: ",main.text0,sep="")
dat1=dat0
mort.plot()



```

```{r CDR.3.age, echo=FALSE}
################################################
age.table=table(dat0$age.group10)
age.groups=names(age.table)
t.max=10

par(mfrow=c(1,1))
for(i in 1:length(age.groups)){
  main.text=paste("Age ",age.groups[i],": ",main.text0,sep="")
  dat1=dat0[dat0$age.group10==age.groups[i],]
  mort.plot()}

```





```{r CDR.3.outcomes, echo=FALSE}
################################################
dat=cdr3.death
table(dat$age.group10)

t1=2
out.2y=ifelse((dat$yrs<=t1)&(dat$event==1),"2-Deceased",
              ifelse(dat$yrs>t1,"0-Alive w/o progression", NA))
table.2y=round(proportions(table(out.2y,dat$age.group10,useNA="ifany"),margin=2),2)
table.2y

t1=5
out.5y=ifelse((dat$yrs<=t1)&(dat$event==1),"2-Deceased",
              ifelse(dat$yrs>t1,"0-Alive w/o progression", NA))

table.5y=round(proportions(table(out.5y,dat$age.group10,useNA="ifany"),margin=2),2)
table.5y

t1=10
out.10y=ifelse((dat$yrs<=t1)&(dat$event==1),"2-Deceased",
              ifelse(dat$yrs>t1,"0-Alive w/o progression", NA))
table.10y=round(proportions(table(out.10y,dat$age.group10,useNA="ifany"),margin=2),2)
table.10y
```

NEW EDIT HERE